---
title: "HW04"
author: "DJames"
date: "October 9, 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## General data reshaping and relationship to aggregation

### Activity 2: Make a tibble with one row per year and columns for life expectancy for two or more countries.

First we need to load the gapminder and tidyverse packages.
```{r}
library(gapminder)
library(tidyverse)
```

We can take a look at the summary of the gapminder data before diving in:
```{r}
summary(gapminder)
str(gapminder)
```

Let's filter for two countries, Brazil and Canada, and then select to only work with the country, year, and life expectancy for each:

```{r}
df_BRCA <- gapminder %>%
  filter(year, country=="Canada" | country=="Brazil") %>% 
  select(year, country, lifeExp)

df_BRCA
```

Let's use the spread function so that each country has its own column with life expectancy listed as the variable of interest underneath, according to year:
```{r}
df_BRCA_long <- df_BRCA %>% 
  spread(key = "country", value = "lifeExp")

df_BRCA_long
#Tried to use knitr::kable() function to make the table more readable but something wasn't working right. Used default instead.
```

Now let's plot the life expectancies for each country against each other over time:
```{r}
ggplot(df_BRCA, aes(year, lifeExp)) +
  geom_line(aes(color=country)) +
  theme_bw()
```

## Join, merge, look up

### Activity :
