---
title: "HW04"
author: "DJames"
date: "October 9, 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## General data reshaping and relationship to aggregation

### Activity 2: Make a tibble with one row per year and columns for life expectancy for two or more countries.

First we need to load the gapminder and tidyverse packages.
```{r}
library(gapminder)
library(tidyverse)
```

We can take a look at the summary of the gapminder data before diving in:
```{r}
summary(gapminder)
str(gapminder)
```

Let's filter for two countries, Brazil and Canada, and then select to only work with the country, year, and life expectancy for each:

```{r}
df_BRCA <- gapminder %>%
  filter(year, country=="Canada" | country=="Brazil") %>% 
  select(year, country, lifeExp)

df_BRCA
```

Let's use the spread function so that each country has its own column with life expectancy listed as the variable of interest underneath, according to year:
```{r}
df_BRCA_long <- df_BRCA %>% 
  spread(key = "country", value = "lifeExp")

df_BRCA_long
#Tried to use knitr::kable() function to make the table more readable but something wasn't working right. Used default instead.
```

Now let's plot the life expectancies for each country against each other over time:
```{r}
ggplot(df_BRCA, aes(year, lifeExp)) +
  geom_line(aes(color=country)) +
  theme_bw()
```

## Join, merge, look up

### Activity 1: Create a second data frame, complementary to Gapminder. Join this with (part of) Gapminder using a  dplyr join function and make some observations about the process and result. Explore the different types of joins:

I'm going to create a data.frame for Brazil's soy harvest over the past two decades, using it as a (bad) proxy for economic growth (which is a component of a country's development) in the country. Then we'll see how it relates to life expectancy over time.
```{r}
df_BRcensus <- "
country,soy harvest (tons),year
Brazil,19897804,1990
Brazil,14937806,1991
Brazil,19214705,1992
Brazil,22590978,1993
Brazil,24931832,1994
Brazil,25682637,1995
Brazil,23166874,1996
Brazil,26392636,1997
Brazil,31307440,1998
Brazil,30987476,1999
Brazil,32820826,2000
Brazil,37907259,2001
Brazil,42107618,2002
Brazil,51919440,2003
Brazil,49549941,2004
Brazil,51182074,2005
Brazil,52464640,2006
Brazil,57857172,2007
Brazil,59833105,2008
Brazil,57345382,2009
Brazil,68756343,2010
Brazil,74815447,2011
Brazil,65848857,2012
Brazil,81724477,2013
Brazil,86760520,2014
"

library(tidyverse)
df_BRcensus <- read_csv(df_BRcensus, trim_ws = TRUE, skip = 1)

```



Now let's join it with the Brazil/Canada subsetted data.frame created in the previous prompt from gapminder:
```{r}
str(df_BRCA)
str(df_BRcensus)
# Need to convert the "year" column for df_BRcensus from character strings to integers
df_BRcensus <- df_BRcensus %>% 
  mutate(year=as.integer(year))
# Join the two data.frames together using inner join, with "year" as the common variable:
df.join <- inner_join(df_BRCA, df_BRcensus, by = "year")
# Filter to get results for only Brazil and delete redundant country column
df.join_filter <- df.join %>% 
  filter(country.x=="Brazil") %>% 
  select(-country.y)
```

Let's gather life expectancy and soy harvest together:
```{r}
df.join_filter_gather <- df.join_filter %>% 
  gather(key = "Measure", value = "Units", lifeExp:`soy harvest (tons)`)
```


Now let's plot the life expecancy of Brazil's population and the soy harvest against time:
```{r}
ggplot(df.join_filter_gather, aes(year, Units)) +
  geom_line(aes(color=Measure)) +
  facet_wrap(~ Measure, scales = "free_y") +
  theme_bw()
```
Both of these measures of development have increased dramatically over the past two decades.
